A Memory-Bounded, Deterministic and Terminating Semantics for the Synchronous Programming Language~\CEU

### 2018-04-15 06-appendix.tex
Cleanup Section 3 and Appendix A up to Theorem 3.8

\begin{definition}
  \label{def.Hnst}
  %%
  A description~$\delta=\<p,n,e>$ is \emph{nested-irre\-ducible}
  iff~$e\ne\nil$ or~$p=\ceu{\Nop},\ceu{\Break}$ or~$\isblocked(p,n)$~is
  true.
\end{definition}

\begin{proposition}
  \label{prop.irr-nst-i}
  %%
  If~$\delta\nst[i]\delta_\Hnst'$ then, for all~$k\ne{i}$, there is
  no~$\delta_\Hnst''$ such that~$\delta\nst[k]\delta''_\Hnst$.
\end{proposition}
\begin{proof}
  By contradiction on the hypothesis that there is such~$k$.
  %%
  Let~$\delta\nst[i]\delta'_\Hnst$, for some~$i\ge0$.
  %%
  There are two cases.
  \begin{case}
    \label{prop.irr-nst-i-case1}
    %%
    Suppose there are~$k>i$ and~$\delta''_\Hnst$ such
    that~$\delta\nst[k]\delta''$.
    %%
    Then, by definition of~$\nst[k]$,
    \begin{equation}
      \label{prop.irr-nst-i-eq1}
      %%
      \delta\nst[i]\delta'\nst[i+1]\delta_1'\nst[i+2]\cdots\nst[k]\delta''.
    \end{equation}
    Since~$\delta'=\<p',n,e'>$ is nested-irreducible, $e'=\nil$
    or~$p=\ceu{\Nop},\ceu{\Break}$ or~$\isblocked(p',n)$.  In any of these
    cases, by the definition of~$\nst$, there is no~$\delta_1'$ such
    that~$\delta'\nst[1]\delta_1'$, which
    contradicts~\eqref{prop.irr-nst-i-eq1}.  Therefore, no such~$k$ can
    exist.
  \end{case}
  \begin{case}
    Suppose there are~$k<i$ and~$\delta''_\Hnst$ such
    that~$\delta\nst[k]\delta''$.  Then, since~$i>k$, by
    Case~\ref{prop.irr-nst-i-case1}, $\delta'$~cannot exist, which is
    absurd.  Therefore, the assumption that there is such~$k$ is
    false.\qedhere
  \end{case}
\end{proof}

\begin{lemma}
  \label{lem.props-nst-i}\strut\\
  %%
  If~$\<p_1,n,e>\nst[i]\<p_1',n,e'>$, for any~$p_2$:
  \begin{enumerate:a}
  \item\label{lem.props-nst-i.a}
    $\<\ceu{p_1;\,p_2},n,e>\nst[i]\<p_1';p_2,n,e'>$.
    %%
  \item\label{lem.props-nst-i.b}
    $\<\ceu{p_1\AtLoop{p_2}},n,e>\nst[i]\<\ceu{p_1'\AtLoop{p_2}},n,e'>$.
    %%
  \item\label{lem.props-nst-i.c}
    $\<\ceu{{p_1}\AtAnd{p_2}},n,e>\nst[i]\<\ceu{{p_1'}\AtAnd{p_2}},n,e'>$.
    %%
  \item\label{lem.props-nst-i.d}
    $\<\ceu{{p_1}\AtOr{p_2}},n,e>\nst[i]\<\ceu{{p_1}'\AtOr{p_2}},n,e'>$.
  \end{enumerate:a}
  \smallskip
  If~$\<p_2,n,e>\nst[i]\<p_2',n,e'>$, for any~$p_1$ such
  that~$\isblocked(p_1,n)$:
  \begin{enumerate:a}
    \setcounter{enumi}{4}
  \item\label{lem.props-nst-i.e}
    $\<\ceu{{p_1}\AtAnd{p_2}},n,e>\nst[i]\<\ceu{{p_1}\AtAnd{p_2'}},n,e'>$.
    %%
  \item\label{lem.props-nst-i.f}
    $\<\ceu{{p_1}\AtOr{p_2}},n,e>\nst[i]\<\ceu{{p_1}\AtOr{p_2'}},n,e'>$.
  \end{enumerate:a}
\end{lemma}
\begin{proof}
  By induction on~$i$.
  %%
  \begin{enumerate:a}
  \item The lemma is trivially true for~$i=0$, as~$p_1=p_1'$, and follows
    directly from~\R{seq-adv} for~$i=1$.  Suppose
    \begin{equation}
      \label{lem.props-nst-i.a.eq1}
      \<p_1,n,e>\nst[1]\<p_1'',n,e''>\nst[i-1]\<p_1',n,e'>\,,
    \end{equation}
    for some~$i>1$.  Then~$\<p_1'',n,e''>$ is not nested-irreducible, i.e.,
    $e=\nil$ and~$p\ne{\ceu{\Nop},\ceu{\Break}}$ and~$\isblocked(p_1'',n)$
    is false.  By~\eqref{lem.props-nst-i.a.eq1} and by~\R{seq-adv},
    \begin{equation}
      \label{lem.props-nst-i.a.eq2}
      \<\ceu{p_1;\,p_2},n,e>\nst[1]\<\ceu{p_1'';\,p_2},n,e''>\,.
    \end{equation}
    From~\eqref{lem.props-nst-i.a.eq1}, by the induction hypothesis,
    \begin{equation}
      \label{lem.props-nst-i.a.eq3}
      \<\ceu{p_1'';\,p_2},n,e''>\nst[i-1]\<\ceu{p_1';\,p_2},n,e'>\,.
    \end{equation}
    From~\eqref{lem.props-nst-i.a.eq2} and~\eqref{lem.props-nst-i.a.eq3},
    \[
      \<\ceu{p_1;\,p_2},n,e>\nst[i]\<\ceu{p_1';\,p_2},n,e'>\,.
    \]

  \item Similar to Case~\eqref{lem.props-nst-i.a}.
    %%
    % The lemma is trivially true for~$n=0$, as~$p_1=p_1'$, and follows
    % directly from~\R{loop-adv} for~$n=1$.  Suppose
    % \begin{equation}
    %   \label{lem.props-nst-i.b.eq1}
    %   \<p_1,n,e>\nst[1]\<p_1'',n,e''>\nst[n-1]\<p_1',n,e'>\,,
    % \end{equation}
    % for some~$n>1$.  Then~$\<p_1'',n,e''>$ is not nested-irreducible.
    % By~\eqref{lem.props-nst-i.b.eq1} and by~\R{loop-adv},
    % \begin{equation}
    %   \label{lem.props-nst-i.b.eq2}
    %   \<\ceu{p_1\AtLoop{p_2}},n,e>\nst[1]\<\ceu{p_1''\Loop{p_2}},n,e''>\,.
    % \end{equation}
    % From~\eqref{lem.props-nst-i.b.eq1}, by the induction hypothesis,
    % \begin{equation}
    %   \label{lem.props-nst-i.b.eq3}
    %   \<\ceu{p_1''\AtLoop{p_2}},n,e''>
    %   \nst[n-1]\<\ceu{p_1'\AtLoop{p_2}},n,e'>\,.
    % \end{equation}
    % From~\eqref{lem.props-nst-i.b.eq2} and~\eqref{lem.props-nst-i.b.eq3},
    % \[
    %   \<\ceu{p_1\AtLoop{p_2}},n,e>\nst[n]\<\ceu{p_1'\AtLoop{p_2}},n,e'>\,.
    % \]
    %%

  \item Similar to Case~\eqref{lem.props-nst-i.a}.
    %%
    % The lemma is trivially true for~$n=0$, as~$p_1=p_1'$, and follows
    % directly from~\R{and-adv1} for~$n=1$.  Suppose
    % \begin{equation}
    %   \label{lem.props-nst-i.c.eq1}
    %   \<p_1,n,e>\nst[1]\<p_1'',n,e''>\nst[n-1]\<p_1',n,e'>\,,
    % \end{equation}
    % for some~$n>1$.  Then~$\<p_1'',n,e''>$ is not nested-irreducible.
    % By~\eqref{lem.props-nst-i.c.eq1} and by~\R{and-adv1},
    % \begin{equation}
    %   \label{lem.props-nst-i.c.eq2}
    %   \<\ceu{{p_1}\AtAnd{p_2}},n,e>
    %   \nst[1]\<\ceu{{p_1}''\AtAnd{p_2}},n,e''>\,.
    % \end{equation}
    % From~\eqref{lem.props-nst-i.c.eq1}, by the induction hypothesis,
    % \begin{equation}
    %   \label{lem.props-nst-i.a.eq3}
    %   \<\ceu{{p_1''}\AtAnd{p_2}},n,e''>
    %   \nst[n-1]\<\ceu{{p_1'}\AtAnd{p_2}},n,e'>\,.
    % \end{equation}
    % From~\eqref{lem.props-nst-i.c.eq2} and~\eqref{lem.props-nst-i.c.eq3},
    % \[
    %   \<\ceu{{p_1}\AtAnd{p_2}},n,e>
    %   \nst[n]\<\ceu{{p_1'}\AtAnd{p_2}},n,e'>\,.
    % \]
    %%

  \item Similar to Case~\eqref{lem.props-nst-i.a}.
    %%
    % The lemma is trivially true for~$n=0$, as~$p_1=p_1'$, and follows
    % directly from~\R{or-adv1} for~$n=1$.  Suppose
    % \begin{equation}
    %   \label{lem.props-nst-i.d.eq1}
    %   \<p_1,n,e>\nst[1]\<p_1'',n,e''>\nst[n-1]\<p_1',n,e'>\,,
    % \end{equation}
    % for some~$n>1$.  Then~$\<p_1'',n,e''>$ is not nested-irreducible.
    % By~\eqref{lem.props-nst-i.d.eq1} and by~\R{or-adv1},
    % \begin{equation}
    %   \label{lem.props-nst-i.d.eq2}
    %   \<\ceu{{p_1}\AtOr{p_2}},n,e>
    %   \nst[1]\<\ceu{{p_1}''\AtOr{p_2}},n,e''>\,.
    % \end{equation}
    % From~\eqref{lem.props-nst-i.d.eq1}, by the induction hypothesis,
    % \begin{equation}
    %   \label{lem.props-nst-i.a.eq3}
    %   \<\ceu{{p_1''}\AtOr{p_2}},n,e''>
    %   \nst[n-1]\<\ceu{{p_1'}\AtOr{p_2}},n,e'>\,.
    % \end{equation}
    % From~\eqref{lem.props-nst-i.d.eq2} and~\eqref{lem.props-nst-i.d.eq3},
    % \[
    %   \<\ceu{{p_1}\AtOr{p_2}},n,e>
    %   \nst[n]\<\ceu{{p_1'}\AtOr{p_2}},n,e'>\,.
    % \]
    %%

  \item The lemma is trivially true for~$i=0$, as~$p_2=p_2'$, and follows
    directly from~\R{and-adv2} for~$i=1$.  Suppose
    \begin{equation}
      \label{lem.props-nst-i.e.eq1}
      \<p_2,n,e>\nst[1]\<p_2'',n,e''>\nst[i-1]\<p_2',n,e'>\,,
    \end{equation}
    for some~$i>1$.  Then~$\<p_2'',n,e''>$ is not nested-irreducible.
    By~\eqref{lem.props-nst-i.e.eq1} and by~\R{and-adv2},
    \begin{equation}
      \label{lem.props-nst-i.e.eq2}
      \<\ceu{{p_1}\AtAnd{p_2}},n,e>
      \nst[1]\<\ceu{{p_1}\AtAnd{p_2''}},n,e''>\,.
    \end{equation}
    From~\eqref{lem.props-nst-i.e.eq1}, by the induction hypothesis,
    \begin{equation}
      \label{lem.props-nst-i.e.eq3}
      \<\ceu{{p_1}\AtAnd{p_2''}},n,e''>
      \nst[i-1]\<\ceu{{p_1}\AtOr{p_2'}},n,e'>\,.
    \end{equation}
    From~\eqref{lem.props-nst-i.e.eq2} and~\eqref{lem.props-nst-i.e.eq3},
    \[
      \<\ceu{{p_1}\AtAnd{p_2}},n,e>
      \nst[i]\<\ceu{{p_1}\AtAnd{p_2'}},n,e'>\,.
    \]

  \item Similar to Case~\eqref{lem.props-nst-i.e}.\qedhere
  \end{enumerate:a}
\end{proof}

\begin{assumption}[Syntactic restrictions]\strut
  \label{ass.syn-rest}
  %%
  \begin{enumerate:a}
  \item\label{ass.syn-rest.fin} If~$p=\ceu{\Fin{p_1}}$ then~$p_1$ contain no
    occurrences of~$\ceu{\AwaitExt}$, $\ceu{\AwaitInt}$, $\ceu{\Every}$,
    or~$\ceu{\Fin}$ expressions.  Consequently, for any~$n$,
    \[
      \<\clear(p_1),n,\nil>\nst[*]\<\ceu{\Nop},n,\nil>\,.
    \]
    %%
  \item\label{ass.syn-rest.loop} If~$p=\ceu{\Loop{p_1}}$ then all execution
    paths of~$p_1$ contain at least one occurrence of a
    matching~$\ceu{\Break}$ or an~$\ceu{\AwaitExt}$ expression.
    Consequently, for all~$n$, there are~$p_1'$ and~$e$ such that
    \[
      \<\ceu{\Loop{p_1}},n,\nil>\nst[*]\<p_1',n,e>\,,
    \]
    where~$p_1'=\ceu{{\Break}\AtLoop{p_1}}$ or~$\isblocked(p_1',n)$\,.
  \end{enumerate:a}
\end{assumption}

\begin{theorem}
  \label{thm.term-nst-*}
  %%
  For any~$\delta$ there is a~$\delta'_\Hnst$ such
  that~$\delta\nst[*]\delta'_\Hnst$.
\end{theorem}
\begin{proof}
  By induction on the structure of programs.
  %%
  Let~$\delta=\<p,n,\nil>$.  The theorem is trivially true if~$\delta$ is
  nested-irreducible, as by definition~$\delta\nst[0]\delta$.  Suppose that
  is not the case.  Then, depending on the structure of~$p$, there are~11
  possibilities.  In each one of them, we show that such~$\delta'_\Hnst$
  indeed exists.
  \begin{case}
    $p=\ceu{\Mem(\Id)}$.
    %%
    Then, by~\R{mem},
    \[
      \<\ceu{\Mem(\Id)},n,\nil>\nst[1]\<\ceu{\Nop},n,\nil>_\Hnst\,.
    \]
  \end{case}

  \begin{case}
    $p=\ceu{\EmitInt(e)}$.
    %%
    Then, by~\R{emit-int},
    \[
      \<\ceu{\EmitInt(e)},n,\nil>\nst[1]\<\ceu{\CanRun(n)},n,e>_\Hnst\,.
    \]
  \end{case}

  \begin{case}
    $p=\ceu{\CanRun(n)}$.
    %%
    Then, by~\R{can-run},
    \[
      \<\ceu{\CanRun(n)},n,\nil>\nst[1]\<\ceu{\Nop},n,\nil>_\Hnst\,.
    \]
  \end{case}

  \begin{case}
    $p=\ceu{\IfElse{\Mem(\Id)}{p'}{p''}}$.
    %%
    There are two subcases.
    \begin{subcase}
      \label{thm.term-nst-*.if-true}
      $\eval(\ceu{\Mem(\Id)})$~is true.
      %%
      Then, by~\R{if-true} and by the induction hypothesis, there is
      a~$\delta'$ such that
      \begin{align*}
        \<\ceu{\IfElse{\Mem(\Id)}{p'}{p''}},n,\nil>
        &\nst[1]\<p',n,e>\\
        &\nst[*]\delta'_\Hnst\,.
      \end{align*}
    \end{subcase}
    \begin{subcase}
      $\eval(\ceu{\Mem(\Id)})$~is false.
      %%
      Similar to Case~\ref{thm.term-nst-*.if-true}.
      %%
      % Then, by~\R{if-false} and by the induction hypothesis, there is
      % a~$\delta'$ such that
      % \begin{align*}
      %   \<\ceu{\IfElse{\Mem(\Id)}{p'}{p''}},n,\nil>
      %   &\nst[1]\<p'',n,e>\\
      %   &\nst[*]\delta'_\Hnst\,.
      % \end{align*}
      %%
    \end{subcase}
  \end{case}

>>>>

\propxirrnsti*
\begin{proof}
  By contradiction on the hypothesis that there is such~$k$.
  %%
  Let~$\delta\nst[i]\delta'_\Hnst$, for some~$i\ge0$.
  %%
  There are two cases.
  \begin{casex}
  \item\label{prop.x.irr-nst-i-case1}
    %%
    Suppose there are~$k>i$ and~$\delta''_\Hnst$ such
    that~$\delta\nst[k]\delta''$.
    %%
    Then, by definition of~$\nst[k]$,
    \begin{equation}
      \label{prop.x.irr-nst-i-eq1}
      %%
      \delta\nst[i]\delta'\nst[i+1]\delta_1'\nst[i+2]\cdots\nst[k]\delta''.
    \end{equation}
    Since~$\delta'=\<p',n,e'>$ is nested-irreducible, $e'=\nil$
    or~$p=\ceu{\Nop},\ceu{\Break}$ or~$\isblocked(p',n)$.  In any of these
    cases, by the definition of~$\nst$, there is no~$\delta_1'$ such
    that~$\delta'\nst[1]\delta_1'$, which
    contradicts~\eqref{prop.x.irr-nst-i-eq1}.  Therefore, no such~$k$ can
    exist.
    %%
  \item Suppose there are~$k<i$ and~$\delta''_\Hnst$ such
    that~$\delta\nst[k]\delta''$.  Then, since~$i>k$, by
    Case~\ref{prop.x.irr-nst-i-case1}, $\delta'$~cannot exist, which is
    absurd.  Therefore, the assumption that there is such~$k$ is
    false.\qedhere
  \end{casex}
\end{proof}

\lemxpropsnsti*
\begin{proof}
  By induction on~$i$.
  %%
  \begin{enumerate:a}
  \item The lemma is trivially true for~$i=0$, as~$p_1=p_1'$, and follows
    directly from~\R{seq-adv} for~$i=1$.  Suppose
    \begin{equation}
      \label{lem.x.props-nst-i.a.eq1}
      \<p_1,n,e>\nst[1]\<p_1'',n,e''>\nst[i-1]\<p_1',n,e'>\,,
    \end{equation}
    for some~$i>1$.  Then~$\<p_1'',n,e''>$ is not nested-irreducible, i.e.,
    $e=\nil$ and~$p\ne{\ceu{\Nop},\ceu{\Break}}$
    and~$\lnot\isblocked(p_1'',n)$.  By~\eqref{lem.x.props-nst-i.a.eq1} and
    by~\R{seq-adv},
    \begin{equation}
      \label{lem.x.props-nst-i.a.eq2}
      \<\ceu{p_1;\,p_2},n,e>\nst[1]\<\ceu{p_1'';\,p_2},n,e''>\,.
    \end{equation}
    From~\eqref{lem.x.props-nst-i.a.eq1}, by the induction hypothesis,
    \begin{equation}
      \label{lem.x.props-nst-i.a.eq3}
      \<\ceu{p_1'';\,p_2},n,e''>\nst[i-1]\<\ceu{p_1';\,p_2},n,e'>\,.
    \end{equation}
    From~\eqref{lem.x.props-nst-i.a.eq2} and~\eqref{lem.x.props-nst-i.a.eq3},
    \[
      \<\ceu{p_1;\,p_2},n,e>\nst[i]\<\ceu{p_1';\,p_2},n,e'>\,.
    \]

  \item Similar to item~\ref{lem.x.props-nst-i.a}.
    %%
    % The lemma is trivially true for~$n=0$, as~$p_1=p_1'$, and follows
    % directly from~\R{loop-adv} for~$n=1$.  Suppose
    % \begin{equation}
    %   \label{lem.x.props-nst-i.b.eq1}
    %   \<p_1,n,e>\nst[1]\<p_1'',n,e''>\nst[n-1]\<p_1',n,e'>\,,
    % \end{equation}
    % for some~$n>1$.  Then~$\<p_1'',n,e''>$ is not nested-irreducible.
    % By~\eqref{lem.x.props-nst-i.b.eq1} and by~\R{loop-adv},
    % \begin{equation}
    %   \label{lem.x.props-nst-i.b.eq2}
    %   \<\ceu{p_1\AtLoop{p_2}},n,e>\nst[1]\<\ceu{p_1''\Loop{p_2}},n,e''>\,.
    % \end{equation}
    % From~\eqref{lem.x.props-nst-i.b.eq1}, by the induction hypothesis,
    % \begin{equation}
    %   \label{lem.x.props-nst-i.b.eq3}
    %   \<\ceu{p_1''\AtLoop{p_2}},n,e''>
    %   \nst[n-1]\<\ceu{p_1'\AtLoop{p_2}},n,e'>\,.
    % \end{equation}
    % From~\eqref{lem.x.props-nst-i.b.eq2} and~\eqref{lem.x.props-nst-i.b.eq3},
    % \[
    %   \<\ceu{p_1\AtLoop{p_2}},n,e>\nst[n]\<\ceu{p_1'\AtLoop{p_2}},n,e'>\,.
    % \]
    %%

  \item Similar to item~\ref{lem.x.props-nst-i.a}.
    %%
    % The lemma is trivially true for~$n=0$, as~$p_1=p_1'$, and follows
    % directly from~\R{and-adv1} for~$n=1$.  Suppose
    % \begin{equation}
    %   \label{lem.x.props-nst-i.c.eq1}
    %   \<p_1,n,e>\nst[1]\<p_1'',n,e''>\nst[n-1]\<p_1',n,e'>\,,
    % \end{equation}
    % for some~$n>1$.  Then~$\<p_1'',n,e''>$ is not nested-irreducible.
    % By~\eqref{lem.x.props-nst-i.c.eq1} and by~\R{and-adv1},
    % \begin{equation}
    %   \label{lem.x.props-nst-i.c.eq2}
    %   \<\ceu{{p_1}\AtAnd{p_2}},n,e>
    %   \nst[1]\<\ceu{{p_1}''\AtAnd{p_2}},n,e''>\,.
    % \end{equation}
    % From~\eqref{lem.x.props-nst-i.c.eq1}, by the induction hypothesis,
    % \begin{equation}
    %   \label{lem.x.props-nst-i.a.eq3}
    %   \<\ceu{{p_1''}\AtAnd{p_2}},n,e''>
    %   \nst[n-1]\<\ceu{{p_1'}\AtAnd{p_2}},n,e'>\,.
    % \end{equation}
    % From~\eqref{lem.x.props-nst-i.c.eq2} and~\eqref{lem.x.props-nst-i.c.eq3},
    % \[
    %   \<\ceu{{p_1}\AtAnd{p_2}},n,e>
    %   \nst[n]\<\ceu{{p_1'}\AtAnd{p_2}},n,e'>\,.
    % \]
    %%

  \item Similar to item~\ref{lem.x.props-nst-i.a}.
    %%
    % The lemma is trivially true for~$n=0$, as~$p_1=p_1'$, and follows
    % directly from~\R{or-adv1} for~$n=1$.  Suppose
    % \begin{equation}
    %   \label{lem.x.props-nst-i.d.eq1}
    %   \<p_1,n,e>\nst[1]\<p_1'',n,e''>\nst[n-1]\<p_1',n,e'>\,,
    % \end{equation}
    % for some~$n>1$.  Then~$\<p_1'',n,e''>$ is not nested-irreducible.
    % By~\eqref{lem.x.props-nst-i.d.eq1} and by~\R{or-adv1},
    % \begin{equation}
    %   \label{lem.x.props-nst-i.d.eq2}
    %   \<\ceu{{p_1}\AtOr{p_2}},n,e>
    %   \nst[1]\<\ceu{{p_1}''\AtOr{p_2}},n,e''>\,.
    % \end{equation}
    % From~\eqref{lem.x.props-nst-i.d.eq1}, by the induction hypothesis,
    % \begin{equation}
    %   \label{lem.x.props-nst-i.a.eq3}
    %   \<\ceu{{p_1''}\AtOr{p_2}},n,e''>
    %   \nst[n-1]\<\ceu{{p_1'}\AtOr{p_2}},n,e'>\,.
    % \end{equation}
    % From~\eqref{lem.x.props-nst-i.d.eq2} and~\eqref{lem.x.props-nst-i.d.eq3},
    % \[
    %   \<\ceu{{p_1}\AtOr{p_2}},n,e>
    %   \nst[n]\<\ceu{{p_1'}\AtOr{p_2}},n,e'>\,.
    % \]
    %%

  \item The lemma is trivially true for~$i=0$, as~$p_2=p_2'$, and follows
    directly from~\R{and-adv2} for~$i=1$.  Suppose
    \begin{equation}
      \label{lem.x.props-nst-i.e.eq1}
      \<p_2,n,e>\nst[1]\<p_2'',n,e''>\nst[i-1]\<p_2',n,e'>\,,
    \end{equation}
    for some~$i>1$.  Then~$\<p_2'',n,e''>$ is not nested-irreducible.
    By~\eqref{lem.x.props-nst-i.e.eq1} and by~\R{and-adv2},
    \begin{equation}
      \label{lem.x.props-nst-i.e.eq2}
      \<\ceu{{p_1}\AtAnd{p_2}},n,e>
      \nst[1]\<\ceu{{p_1}\AtAnd{p_2''}},n,e''>\,.
    \end{equation}
    From~\eqref{lem.x.props-nst-i.e.eq1}, by the induction hypothesis,
    \begin{equation}
      \label{lem.x.props-nst-i.e.eq3}
      \<\ceu{{p_1}\AtAnd{p_2''}},n,e''>
      \nst[i-1]\<\ceu{{p_1}\AtOr{p_2'}},n,e'>\,.
    \end{equation}
    From~\eqref{lem.x.props-nst-i.e.eq2} and~\eqref{lem.x.props-nst-i.e.eq3},
    \[
      \<\ceu{{p_1}\AtAnd{p_2}},n,e>
      \nst[i]\<\ceu{{p_1}\AtAnd{p_2'}},n,e'>\,.
    \]

  \item Similar to item~\ref{lem.x.props-nst-i.e}.\qedhere
  \end{enumerate:a}
\end{proof}

\thmxtermnstx*
\begin{proof}
  By induction on the structure of programs.
  %%
  Let~$\delta=\<p,n,\nil>$.  The theorem is trivially true if~$\delta$ is
  nested-irreducible, as by definition~$\delta\nst[0]\delta$.  Suppose that
  is not the case.  Then, depending on the structure of~$p$, there are~11
  possibilities.  In each one of them, we show that such~$\delta'_\Hnst$
  indeed exists.
  \begin{casex}
  \item $p=\ceu{\Mem(\Id)}$.
    %%
    Then, by~\R{mem},
    \[
      \<\ceu{\Mem(\Id)},n,\nil>\nst[1]\<\ceu{\Nop},n,\nil>_\Hnst\,.
    \]

  \item $p=\ceu{\EmitInt(e)}$.
    %%
    Then, by~\R{emit-int},
    \[
      \<\ceu{\EmitInt(e)},n,\nil>\nst[1]\<\ceu{\CanRun(n)},n,e>_\Hnst\,.
    \]

  \item $p=\ceu{\CanRun(n)}$.
    %%
    Then, by~\R{can-run},
    \[
      \<\ceu{\CanRun(n)},n,\nil>\nst[1]\<\ceu{\Nop},n,\nil>_\Hnst\,.
    \]

  \item $p=\ceu{\IfElse{\Mem(\Id)}{p'}{p''}}$.
    %%
    There are two subcases.
    \begin{casex}
    \item\label{thm.x.term-nst-*.if-true} $\eval(\ceu{\Mem(\Id)})$.
      %%
      Then, by~\R{if-true} and by the induction hypothesis, there is
      a~$\delta'$ such that
      \begin{align*}
        \<\ceu{\IfElse{\Mem(\Id)}{p'}{p''}},n,\nil>
        &\nst[1]\<p',n,e>\\
        &\nst[*]\delta'_\Hnst\,.
      \end{align*}
      %%
    \item$\lnot\eval(\ceu{\Mem(\Id)})$.
      %%
      Similar to Case~\ref{thm.x.term-nst-*.if-true}.
      %%
      % Then, by~\R{if-false} and by the induction hypothesis, there is
      % a~$\delta'$ such that
      % \begin{align*}
      %   \<\ceu{\IfElse{\Mem(\Id)}{p'}{p''}},n,\nil>
      %   &\nst[1]\<p'',n,e>\\
      %   &\nst[*]\delta'_\Hnst\,.
      % \end{align*}
      %%
    \end{casex}

  \item $p=\ceu{p';\,p''}$.
    %%
    There are three subcases.
    \begin{casex}
    \item\label{thm.x.term-nst-*.seq-nop} $p'=\ceu{\Nop}$.
      %%
      Then, by~\R{seq-nop} and by the induction hypothesis, there is
      a~$\delta'$ such that
      \[
        \<\ceu{\Nop;\,p''},n,\nil>
        \nst[1]\<p'',n,e>\nst[*]\delta'_\Hnst\,.
      \]
      %%
    \item\label{thm.x.term-nst-*.seq-brk} $p'=\ceu{\Break}$.
      %%
      Then, by~\R{seq-brk},
      \[
        \<\ceu{\Break;\,p''},n,\nil>\nst[1]\<\ceu{\Break},n,\nil>_\Hnst\,.
      \]
      %%
    \item\label{thm.x.term-nst-*.seq-adv}
      $p'\ne\ceu{\Nop},\ceu{\Break}$.
      %%
      Then, by the induction hypothesis, there are~$p_1'$ and~$e$ such that
      \[
        \<p',n,\nil>\nst[*]\<p_1',n,e>_\Hnst\,.
      \]
      By item~\ref{lem.x.props-nst-i.a} of Lemma~\ref{lem.x.props-nst-i},
      \begin{equation}
        \label{thm.x.term-nst-*.seq-adv.eq1}
        \<\ceu{p';\,p''},n,\nil>\nst[*]\<\ceu{p_1';\,p''},n,e>\,.
      \end{equation}
      It remains to be shown that~$\<\ceu{p_1';\,p''},n,e>$ is
      nested-irreducible.  There are four possibilities following from the
      fact that the simpler~$\<p_1',n,e>$ is nested-irreducible.
      %%
      \begin{casex}
      \item $e\ne\nil$.  Then, by the definition of~$\Hnst$,
        description~$\<\ceu{p_1';\,p''},n,e>$ is nested-irreducible.
        %%
      \item $p_1'=\ceu{\Nop}$.
        %%
        From~\eqref{thm.x.term-nst-*.seq-adv.eq1},
        \[
          \<\ceu{p';\,p''},n,\nil>\nst[*]\<\ceu{\Nop;\,p''},n,e>\,.
        \]
        From this point on, this case is similar to
        Case~\ref{thm.x.term-nst-*.seq-nop}.
        %%
      \item $p_1'=\ceu{\Break}$.
        %%
        From~\eqref{thm.x.term-nst-*.seq-adv.eq1},
        \[
          \<\ceu{p';\,p''},n,\nil>\nst[*]\<\ceu{\Break;\,p''},n,e>\,.
        \]
        From this point on, this case is similar to
        Case~\ref{thm.x.term-nst-*.seq-brk}.
        %%
      \item$\isblocked(p_1',n)$.
        %%
        Then, by definition,
        \[
          \isblocked(\ceu{p_1';p''},n)=\isblocked(p_1',n)=\mathit{true}\,.
        \]
        Hence, from~\eqref{thm.x.term-nst-*.seq-adv.eq1} and by the
        definition~$\Hnst$, description~$\<\ceu{p_1';\,p''},n,e>$ is
        nested-irreducible.
      \end{casex}
    \end{casex}

  \item\label{thm.x.term-nst-*.loop}
    $p=\ceu{\Loop{p'}}$.
    %%
    Then, by item~\ref{ass.x.syn-rest.loop} of
    Assumption~\ref{ass.x.syn-rest},
    \begin{equation}\label{thm.x.term-nst-*.loop-expd.eq1}
      \<\ceu{\Loop{p'}},n,\nil>\nst[*]\<p_1',n,e>\,,
    \end{equation}
    for some~$e$ and~$p_1'$ such that either~$p_1'=\ceu{\Break\AtLoop{p'}}$
    or~$\isblocked(p_1',n)$.
    \begin{casex}
    \item$p_1'=\ceu{\Break\AtLoop{p'}}$.
      %%
      From~\eqref{thm.x.term-nst-*.loop-expd.eq1}, by~\R{loop-brk},
      \begin{align*}
        \<\ceu{\Loop{p'}},n,\nil>
        &\nst[*]\<\ceu{\Break\AtLoop{p'}},n,e>\\
        &\nst[1]\<\ceu{\Nop},n,e>_\Hnst\,.
      \end{align*}
      %%
    \item$\isblocked(p_1',n)$.  Hence,
      from~\eqref{thm.x.term-nst-*.loop-expd.eq1} and by the definition
      of~$\Hnst$, $\<p_1',n,e>_\Hnst$.
    \end{casex}

  \item$p=\ceu{p'\AtLoop{p''}}$.
    %%
    There are three subcases.
    \begin{casex}
    \item$p'=\ceu{\Nop}$.
      %%
      Then, by~\R{loop-nop},
      \[
        \<\ceu{\Nop\AtLoop{p''}},n,\nil>
        \nst[1]\<\ceu{\Loop{p''}},n,\nil>\,.
      \]
      From this point on, this case is similar to
      Case~\ref{thm.x.term-nst-*.loop}.
      %%
    \item$p'=\ceu{\Break}$.  Then, by~\R{loop-brk},
      \[
        \<\ceu{\Break\AtLoop{p''}},n,\nil>
        \nst[1]\<\ceu{\Nop},n,\nil>_\Hnst\,.
      \]
      %%
    \item$p'\ne\ceu{\Nop},\ceu{\Break}$.  Then, by the induction hypothesis,
      there are~$p_1'$ and~$e$ such that
      \[
        \<p',n,\nil>\nst[*]\<p'_1,n,e>_\Hnst\,.
      \]
      By item~\ref{lem.x.props-nst-i.b} of Lemma~\ref{lem.x.props-nst-i},
      \[
        \<\ceu{p'\AtLoop{p''}},n,\nil>
        \nst[*]\<\ceu{p_1'\AtLoop{p''}},n,e>\,.
      \]
      It remains to be show that~$\<\ceu{p_1'\AtLoop{p''}},n,e>$ is
      nested-irreducible.  The rest of this proof is similar to that of
      Case~\ref{thm.x.term-nst-*.seq-adv}.
    \end{casex}

  \item$p=\ceu{{p'}\And{p''}}$.
    Then, by~\R{and-expd},
      \<\ceu{{p'}\And{p''}},n,\nil>
      \nst[1]\<\ceu{{p'}\AtAnd{(\CanRun(n);\,p'')}},n,\nil>\,.
    Case~\ref{thm.x.term-nst-*.and}.
  \item\label{thm.x.term-nst-*.and}
